---
title: "Analysis"
---

First we load the packages:

```{r}
#| output: false

library(babynames)
library(knitr)
library(dplyr)
library(ggplot2)
```

The first ten lines of the babynames dataset looks like:

```{r}
head(babynames) |> kable()
```

To create a visualization of the most popular baby names, we have created two functions. Click the 'Code' link to view:

```{r}
#| code-fold: true
get_most_frequent <- function(babynames, select_sex, from = 1950) {
  most_freq <- babynames |>
    filter(sex == select_sex, year > from) |>
    group_by(name) |>
    summarise(average = mean(prop)) |>
    arrange(desc(average))
  return(most_freq)
}

plot_top <- function(x, babynames, select_sex, top = 10, from = 1950) {
  topx <- x$name[1:top]
  
  p <- babynames |>
    filter(name %in% topx, sex == select_sex, year > from) |>
    ggplot(aes(x = year, y = prop, color = name)) +
    geom_line() +
    scale_color_brewer(palette = "Paired") +
    theme_classic()
  
  return(p)
}
```

Here we call the code to visualize the top 10 most frequent girl names from 1950 onwards:

```{r}
#| label: fig-line-girls
#| fig-cap: "Line plot displaying proportion of top 10 girl names by year"
get_most_frequent(babynames, select_sex = "F") |>
  plot_top(babynames, select_sex = "F", top = 10)
```

In @fig-line-girls you can see that there has been a peak in popularity for 'Lisa', 'Jennifer' and 'Jessica'. Interesting! Let's have a look at the boys names:


```{r}
#| label: fig-line-boys
#| fig-cap: "Line plot displaying proportion of top 10 boy names by year"
most_freq_boys <- get_most_frequent(babynames, select_sex = "M")
plot_top(most_freq_boys, babynames, select_sex = "M", top = 10)
```

@fig-line-boys shows that names that were popular before 1990 are relatively infrequent after 2000. 